---
title: "Figures and calculations"
format: html
editor: visual
---

## Libraries

```{r, output = FALSE, message = FALSE}

library(tidyverse)
library(ggvenn)
library(here)
library(grDevices)
library(RaMS)
library(ggpointdensity)
library(patchwork)
library(readxl)
library(gghalves)
library(igraph)
library(ggforce)
library(ggraph)

```

## Colors

```{r}

plot_palette <- function(palette) {
  g <- ggplot2::ggplot(
    data = data.frame(
      x = seq_len(length(palette)),
      y = "1",
      fill = palette
    ),
    mapping = ggplot2::aes(
      x = x, y = y, fill = fill
    )
  ) +
    ggplot2::geom_tile() +
    ggplot2::scale_fill_identity() +
    ggplot2::theme_void()
  return(g)
}

choose_colours <- c("navy","cornflowerblue","purple", "hotpink", "orange", "forestgreen")
all_colours <- colorRampPalette(choose_colours)(100)
plot_palette(all_colours) / plot_palette(choose_colours)

inq_dark_blue <- choose_colours[1]
inq_blue <- choose_colours[2]
inq_purple <- choose_colours[3]
inq_pink <- choose_colours[4]
inq_orange <- choose_colours[5]
inq_green <- choose_colours[6]
```

## Data

### Mixed proteome

```{r}

mixed_psms <- 
  read.csv('../Results/mixed_proteome/mixed_proteome_normal_unfiltered_psm_table.csv')
mixed_peptide <- read.csv('../Results/mixed_proteome/mixed_proteome_normal_unfiltered_peptide_table.csv')


mixed_MBR_psms <- read.csv('../Results/mixed_proteome/mixed_proteome_MBR_unfiltered_psm_table.csv')

mixed_MBR_peptide <- read.csv('../Results/mixed_proteome/mixed_proteome_MBR_unfiltered_peptide_table.csv')

mixed_MBR_proteins <- read.csv('../Results/mixed_proteome/mixed_proteome_MBR_unfiltered_protein_table.csv')


mixed_preds <- 
  read.csv('../Datafiles/mixed_proteome/mixed.csv') |>
  mutate(file_ID = str_split_i(file, '\\_',-1)) |>
  mutate(scan_number = scan_number-1)

mixed_MBR_preds <- 
  read.csv('../Datafiles/mixed_proteome/mixed_proteome_MBR_preds.csv') |>
  mutate(file_ID = str_split_i(file, '\\_',-1)) 


mixed_database_proteins <- 
  read_xlsx('../Datafiles/mixed_proteome/Mixed_proteome_proteins.xlsx') |>
   mutate(principal_protein = str_split_i(Accession, '\\s',1))

mixed_database_peptides <- 
  read_xlsx('../Datafiles/mixed_proteome/Mixed_proteome_peptides.xlsx')

mixed_database_psms <- 
  read_xlsx('../Datafiles/mixed_proteome/Mixed_proteome_psms.xlsx')


file_ids_mixed_database <- mixed_database_psms |>
  mutate(file_id = `File ID`,
         file_name = `Spectrum File`) |>
  distinct() |>
  mutate(our_file_id = str_split_i(str_split_i(file_name, '\\.',1),'\\_',-1)) |>
  select(file_id, our_file_id) |>
  unique()

replace_names <- setNames(file_ids_mixed_database$our_file_id, file_ids_mixed_database$file_id)

mixed_database_psms <- mixed_database_psms |>
  mutate(`File ID` = str_split_i(str_split_i(`Spectrum File`, '\\.',1),'\\_',-1))

mixed_database_peptides <- mixed_database_peptides |>
  rename_with(~ str_replace_all(.x, replace_names), contains("Abundance"))

mixed_fasta <- tibble(full = names(
  seqinr::read.fasta('../Datafiles/mixed_proteome/mixed_human_yeast_ecoli.fasta', 
                     as.string = TRUE))) |>
  mutate(accession = str_split_i(full, '\\|', 2),
         species = str_split_i(full, '\\_', 2)) |>
  select(accession, species)

mixed_database_peptides <- mixed_database_peptides |> 
  mutate(principal_protein = str_split_i(`Master Protein Accessions`,';',1)) |>
  mutate(species = mixed_fasta$species[match(principal_protein, mixed_fasta$accession)])
```

#### Experiment files

```{r}
ids <- c('A1','A2','A3','B1','B2','B3')

paths <- str_c('../Datafiles/mixed_proteome/mzml/20240221_MK_MEKONG_KK_FAIMS_1CV_Evo_40SPD_Whisper_1475_mixed_proteome_',ids,'.mzML')

for (file in paths){
  id = str_split_i(str_split_i(file, '\\_', -1), '\\.', 1)
  
  mzML <- grabMzmlData(file, c('MS1','MS2'))
  
  assign(str_c('mixed_specs_',id), 
    mzML$MS1 |> 
    group_by(rt) |>
    summarise(mz = list(mz), int = list(int), .groups = 'drop') |>
    bind_rows(mzML$MS2 |>
              select(rt, premz) |>
              distinct()) |>
    arrange(rt) |>
    mutate(scan_nr = cur_group_rows()) |>
    mutate(scan_nr = scan_nr-1) |>
    mutate(MSLevel = ifelse(is.na(premz), '1','2')))
}

```

### GluC

```{r}

gluc_psms <- read.csv('../Results/GluC/Gluc_normal_unfiltered_psm_table.csv')
gluc_peptides <- read.csv('../Results/GluC/Gluc_normal_unfiltered_peptide_table.csv')
gluc_proteins <- read.csv('../Results/GluC/Gluc_normal_unfiltered_protein_table.csv')

gluc_MBR_psms <- read.csv('../Results/GluC/GluC_MBR_unfiltered_psm_table.csv')
gluc_MBR_peptides <- read.csv('../Results/GluC/Gluc_MBR_unfiltered_peptide_table.csv')
gluc_MBR_proteins <- read.csv('../Results/GluC/Gluc_MBR_unfiltered_protein_table.csv')

gluc_database_psms <- read_xlsx('../Datafiles/GluC/GluC_search_PSMs.xlsx')
gluc_database_peptides <- read_xlsx('../Datafiles/GluC/GluC_search_peptides.xlsx')
gluc_database_proteins <- read_xlsx('../Datafiles/GluC/GluC_search_proteins.xlsx')

file_ids_gluc_database <- gluc_database_psms |>
  mutate(file_id = `File ID`,
         file_name = `Spectrum File`) |>
  distinct() |>
  mutate(our_file_id = str_split_i(str_split_i(file_name, '\\.',1),'\\_',-1)) |>
  select(file_id, our_file_id) |>
  unique()

replace_names <- setNames(file_ids_gluc_database$our_file_id, file_ids_gluc_database$file_id)

gluc_database_psms <- gluc_database_psms |>
  mutate(`File ID` = str_split_i(str_split_i(`Spectrum File`, '\\.',1),'\\_',-1))

gluc_database_peptides <- gluc_database_peptides |>
  rename_with(~ str_replace_all(.x, replace_names), contains("Abundance"))

gluc_database_proteins <- gluc_database_proteins |>
  rename_with(~ str_replace_all(.x, replace_names), contains("Abundance"))
```

```{r}

gluc_preds <- 
  read_csv('../Datafiles/GluC/gluc_labelled_kpreds.csv',
    show_col_types = FALSE) |> 
  mutate(file = str_split_i(str_split_i(sample, "_", -1), "\\.", 1), 
         probs = 10**log_probs) |> 
  rename(scan_number = spectrum_index,
         target = targets, 
         log_probs_preds = log_probs, 
         probs_preds = probs) |> 
  mutate(target = str_replace_all(target, "I", "L"), 
         preds = str_replace_all(preds, "I", "L")) 

preds_train <-
  read_csv('../Results/GluC/GluC_calibrated_predictions.csv',
    show_col_types = FALSE) |> 
  mutate(file = str_split_i(str_split_i(file, "_", -1), "\\.", 1)) |>
  filter(train == FALSE) |> 
  mutate(ppm_cali = calibrated_mz_error * 1e6 / theo_mz, 
         ppm_iso = iso_adj_mz_error * 1e6 / theo_mz)

gluc_calibrated_preds <- gluc_preds |> 
  full_join(preds_train, by = c("file", "scan_number", "preds")) |> 
  mutate(correct = ifelse(preds == target, TRUE, FALSE)) 

gluc_recalibrated_preds <- 
  read_csv('../Results/GluC/GluC_recalibration_predictions.csv',
    show_col_types = FALSE) |> 
  mutate(file = str_split_i(experiment_name, "_", 1), 
         probs = 10**log_probs) |> 
  rename(log_probs_cali = log_probs, 
         probs_cali = probs, 
         preds_cali = preds) |> 
  mutate(preds_cali = str_replace_all(preds_cali, "I", "L"))

```

### NPv3

```{r}

NPv3_psms <- read.csv('../Results/NPv3/NPv3_normal_unfiltered_psm_table.csv')

NPv3_MBR_psms <- read.csv('../Results/NPv3/NPv3_MBR_unfiltered_psm_table.csv')
NPv3_MBR_peptide <- read.csv('../Results/NPv3/NPv3_MBR_unfiltered_peptide_table.csv')
NPv3_MBR_proteins <- read.csv('../Results/NPv3/NPv3_MBR_unfiltered_protein_table.csv')

NPv3_MBR_proteins_ungrouped <- read.csv('../Results/NPv3/NPv3_MBR_unfiltered_protein_table_ungrouped.csv')

```

## Plot functions

### Save plots

```{r}

save_fig <- function(plot, name, show = TRUE, w = 1920, h = 1440) {
  path <- str_c(getwd(), '/', name)
  ggsave(filename = path, 
         plot = plot, 
         width = w, 
         height = h, 
         units = 'px',
         dpi = 300,
         bg = 'white')
  if (show){
    show(plot)
  }
}

```

### Plot quantification area with RT, MS1 intensities, identification and boundaries

```{r}

plot_quantification_area <- function(specs, preds, target_seq_spec, fileID) {

  if (str_detect(target_seq_spec,'[A-Z]')) {
    target_seq <- target_seq_spec
    target_spec <- preds |> 
      filter(file_ID == fileID) |>
      filter(preds == target_seq) |>
      arrange(desc(10^log_probs)) |>
      head(1) |>
      pull(scan_number)
  }
  else {
    target_spec <- target_seq_spec
    target_seq <- preds |>
      filter(file_ID == fileID) |>
      filter(scan_number == target_spec) |>
      pull(preds)
  }
  
  conf <- preds |>
    filter(file_ID == fileID) |>
    filter(scan_number == target_spec) |>
    pull(log_probs)
  
  conf <- 10^conf
  
  target_RT <- specs$rt[match(target_spec, specs$scan_nr)]
  tolerance_RT <- 1/100 * target_RT
  
  target_MZ <- specs$premz[match(target_spec, specs$scan_nr)]
  
  tolerance_ppm <- 97.63457296020923 # Mixed proteome tolerance
  tolerance <- (target_MZ*tolerance_ppm)/10**6
  
  data <- specs |>
    filter(MSLevel == 1) |>
    unnest(cols = c(mz,int)) |>
    filter(abs(rt-target_RT) <= tolerance_RT) |>
    filter(abs(mz-target_MZ) <= tolerance) |>
    group_by(rt) |>
    summarise(int = max(int)) 
  
  int_target <- data |>
    filter(rt <= target_RT) |>
    arrange(desc(rt)) |>
    head(1) |>
    pull(int)
  
  min_bound <- data |>
    filter(rt <= target_RT) |>
    arrange(desc(rt)) |> 
    filter(lead(int) < int_target*0.2) |> 
    filter(int < int_target*0.2) |>
    head(1) |>
    pull(rt)
  
  max_bound <- data |>
    filter(rt >= target_RT) |>
    arrange(rt) |> 
    filter(lead(int) < int_target*0.2) |> 
    filter(int < int_target*0.2) |> 
    head(1) |>
    pull(rt)
  
  if (length(max_bound) == 0){
    max_bound <- max(data |> pull(rt))
  }
  if (length(min_bound) == 0){
    min_bound <- min(data |> pull(rt))
  }
  
  p <- ggplot(data, aes(x = rt, y = int)) +
        geom_line(color = inq_dark_blue, show.legend = FALSE) + 
        geom_ribbon(data = subset(data, rt >= min_bound & rt <= max_bound),
                    aes(ymin = 0, ymax = int, color = "Quantification area"),
                    alpha = 0.3, fill = inq_blue, linewidth = 0) +
        geom_vline(aes(xintercept = target_RT, color = "Identification scan"), 
                   linetype = "solid", size = 0.8) + 
        geom_vline(aes(xintercept = min_bound, color = "Boundary"), 
                   linetype = "dashed", size = 0.8) +
        geom_vline(aes(xintercept = max_bound, color = "Boundary"), 
                   linetype = "dashed", size = 0.8) +
        geom_hline(aes(yintercept = int_target * 0.2, color = "20% intensity threshold"), 
                   linetype = "dotted", size = 1) +
        scale_color_manual(
          name = NULL,
          values = c(
            "Identification scan" = inq_pink,
            "Boundary" = inq_green,
            "20% intensity threshold" = inq_purple,
            "Quantification area" = inq_blue
          )
        ) +
        labs(
          subtitle = str_c('File: ',
                           fileID,
                           ' | RT window: +/-', 
                           round(tolerance_RT*60, digits = 2), 
                           ' seconds',
                           ' | Scan number: ', 
                           target_spec, 
                           ' | Confidence: ',
                           round(conf, digits = 2)),
          x = 'RT [min]',
          y = 'Intensity'
        ) +
        theme_minimal() +
        theme(
          legend.position = c(0.98, 0.98),  
          legend.justification = c("right", "top"),
          legend.background = element_rect(fill = "white", color = "white"),
          legend.box.background = element_rect(color = "white"),
          legend.key.height = unit(0.7, "cm"),
          legend.spacing.y = unit(-0.3, "cm"),
          plot.background = element_rect(color = 'white')
        ) +
        guides(
          color = guide_legend(
            override.aes = list(
              size = 3  
            )
          )
        )

return(p)
}
```

### Make peak comparison for MBR sequences

```{r}

get_plot_data <- function(target_MZ, target_RT, ID, return_RT){
  
  target_RT = target_RT / 60
  
  tolerance_RT <- 2/100 * target_RT

  tolerance_ppm <- 97.63457296020923 # Mixed proteome tolerance
  tolerance <- (target_MZ*tolerance_ppm)/10**6
  
  data <- get(str_c('mixed_specs_', ID)) |>
    filter(MSLevel == 1) |>
    unnest(cols = c(mz,int)) |>
    filter(abs(rt-target_RT) <= tolerance_RT) |>
    filter(abs(mz-target_MZ) <= tolerance) |>
    group_by(rt) |>
    summarise(int = max(int)) 
  
  ints <- data |> pull(int)
  rts <- data |> pull(rt)
  
  if (return_RT){
    return(list(rts))
  }
  else {
    return(list(ints))
  }
}


plot_MBR_comparison <- function(sequence, next_group = 1, exp_order = NULL) {

  if (!is.null(exp_order)) {
    exp_order <- exp_order
  } else {
    exp_order <- c('A1','A2','A3','B1','B2','B3')
  }
  
  data <- 
    mixed_MBR_preds |>
    select(-file) |>
    group_by(preds) |>
    filter(n() >= 6) |>
    mutate(group_min_rt = min(rt),
           group_max_rt = max(rt)) |>
    arrange(group_min_rt) |>
    filter(preds == sequence | !any(is.na(log_probs)) & n() == 6 & n_distinct(file_ID) == 6) |>
    ungroup() |>
    group_by(group_min_rt)
  
  sequence_data <- data |> filter(preds == sequence)
  seq_min_rt <- min(sequence_data$group_min_rt)
  seq_mz <- str_c(sequence,', m/z: ', round(head(sequence_data, 1)$mz, digits = 2))
  
  data <- data |>
    filter(group_max_rt < seq_min_rt | preds == sequence) |>
    mutate(group_id = cur_group_id()) 
  
  seq_group_id <- data |> filter(preds == sequence) |> head(1) |> pull(group_id)
  
  data <- data |>
    filter(preds == sequence  | group_id == seq_group_id-next_group)
  
  prior_seq_mz <- data |> filter(preds != sequence) |> head(1) |> 
    mutate(name = str_c(preds,', m/z: ', round(mz, digits = 2))) |> pull(name)
  
  color_values = setNames(
                  c(inq_green, inq_blue, inq_orange, inq_pink),
                  c(prior_seq_mz, seq_mz, 'Adjusted identification', 'Identification')
                )
  
  p <- data |>
    ungroup() |>
    select(-group_min_rt, -group_max_rt) |>
    rowwise() |>
    mutate(
      rts = get_plot_data(target_MZ = mz, target_RT = rt, ID = file_ID, return_RT = TRUE),
      ints = get_plot_data(target_MZ = mz, target_RT = rt, ID = file_ID, return_RT = FALSE),
      mbr = ifelse(is.na(log_probs), TRUE, FALSE),
      preds = str_c(preds,', m/z: ', round(mz, digits = 2)),
      file_ID = factor(file_ID, levels = exp_order)
    ) |>
    mutate(true_rt = rts[which.max(ints)]) |>
    unnest(cols = c(rts, ints)) |>
    ggplot(aes(x = rts, y = ints, color = preds)) +
      geom_line() + 
      geom_vline(aes(xintercept = rt / 60, linetype = mbr, color = 'Identification')) +  
      geom_vline(data = function(d) d |> filter(mbr == TRUE), 
                 aes(xintercept = true_rt, color = 'Adjusted identification'), 
                 linetype = 'dotdash') + 
      scale_color_manual(values = color_values) + 
      theme_minimal() +
      facet_grid(file_ID ~ ., scales = 'free') +
      labs(
        title = "Reference and target sequence MBR example",
        x = "Retention Time (minutes)",
        y = "Intensity",
        linetype = "MBR identification",
        color = NULL
      )
  
  return(p)
}

```

### Plot quantification area for MBR with RT, MS1 intensities, initial and final identification and boundaries

```{r}

plot_quantification_area_MBR_full <- function(specs, seq, target_MZ, initial_RT, fileID) {
  
  initial_RT <- initial_RT/60
  
  initial_tolerance <- 1.01*1.02 * initial_RT
  
  tolerance_ppm <- 97.63457296020923 # Mixed proteome tolerance
  tolerance <- (target_MZ*tolerance_ppm)/10**6
  
  data <- specs |>
    filter(MSLevel == 1) |>
    unnest(cols = c(mz,int)) |>
    filter(abs(rt-initial_RT) <= initial_tolerance) |>
    filter(abs(mz-target_MZ) <= tolerance) |>
    group_by(rt) |>
    summarise(int = max(int)) 
  
  subset_data <- data |>
    filter(abs(rt-initial_RT) <= 2/100 * initial_RT) 
  
  subset_min <- min(subset_data$rt)
  subset_max <- max(subset_data$rt)
  
  final_RT <- subset_data |>
    arrange(desc(int)) |>
    head(1) |>
    pull(rt)
  
  int_target <- subset_data |>
    arrange(desc(int)) |>
    head(1) |>
    pull(int)
  
  final_tolerance <- 1/100 * final_RT
  
  final_data <- data |>
    filter(abs(rt-final_RT) <= final_tolerance)
  
  min_bound <- data |>
    filter(rt <= final_RT) |>
    arrange(desc(rt)) |> 
    filter(lead(int) < int_target*0.2) |> 
    filter(int < int_target*0.2) |>
    head(1) |>
    pull(rt)
  
  max_bound <- data |>
    filter(rt >= final_RT) |>
    arrange(rt) |> 
    filter(lead(int) < int_target*0.2) |> 
    filter(int < int_target*0.2) |> 
    head(1) |>
    pull(rt)
  
  plot_data <- data |>
    filter(rt >= min(subset_min, min_bound)) |>
    filter(rt <= max(subset_max, max_bound))
  
  p <- ggplot(plot_data, aes(x = rt, y = int)) +
        geom_line(color = inq_dark_blue, show.legend = FALSE) + 
        geom_ribbon(data = subset(data, rt >= min_bound & rt <= max_bound),
                    aes(ymin = 0, ymax = int, color = "Quantification area"),
                    alpha = 0.3, fill = inq_blue, linewidth = 0) +
        geom_vline(aes(xintercept = initial_RT, color = "Initial identification"), 
                   linetype = "dotted", size = 0.8) + 
        geom_vline(aes(xintercept = final_RT, color = "Final identification"), 
                   linetype = "solid", size = 0.8) + 
        geom_vline(aes(xintercept = subset_min, color = "Search boundary"), 
                   linetype = "dotdash", size = 0.8) +
        geom_vline(aes(xintercept = subset_max, color = "Search boundary"), 
                   linetype = "dotdash", size = 0.8) +
        geom_vline(aes(xintercept = min_bound, color = "Boundary"), 
                   linetype = "dashed", size = 0.8) +
        geom_vline(aes(xintercept = max_bound, color = "Boundary"), 
                   linetype = "dashed", size = 0.8) +
        geom_hline(aes(yintercept = int_target * 0.2, color = "20% intensity threshold"), 
                   linetype = "dotted", size = 1) +
        scale_color_manual(
          name = NULL,
          values = c(
            "Initial identification" = inq_orange,
            "Final identification" = inq_pink,
            "Search boundary" = inq_blue,
            "Boundary" = inq_green,
            "20% intensity threshold" = inq_purple,
            "Quantification area" = inq_blue
          )
        ) +
        labs(
          subtitle = str_c('File: ',
                           fileID,
                           ' | Initial RT: ', 
                           round(initial_RT, digits = 2), 
                           ' min',
                           ' | Final RT: ', 
                           round(final_RT, digits = 2), 
                           ' min',
                           ' | Final RT window: +/-',
                           round(final_RT*0.01*60, digits = 2),
                           ' seconds'),
          x = 'RT [min]',
          y = 'Intensity'
        ) +
        theme_minimal() +
        theme(
          legend.justification = c("right"),
          legend.background = element_rect(fill = "white", color = "white"),
          legend.box.background = element_rect(color = "white"),
          legend.key.height = unit(0.7, "cm"),
          legend.spacing.y = unit(-0.3, "cm"),
          plot.background = element_rect(color = 'white')
        ) +
        guides(
          color = guide_legend(
            override.aes = list(
              size = 3  
            )))
  
  return(p)
  
}
```

### Plot quantification area for MBR with RT, MS1 intensities, final identification and boundaries

```{r}

plot_quantification_area_MBR <- function(specs, seq, target_MZ, initial_RT, fileID) {
  
  initial_RT <- initial_RT/60
  
  initial_tolerance <- 1.01*1.02 * initial_RT
  
  tolerance_ppm <- 97.63457296020923 # Mixed proteome tolerance
  tolerance <- (target_MZ*tolerance_ppm)/10**6
  
  data <- specs |>
    filter(MSLevel == 1) |>
    unnest(cols = c(mz,int)) |>
    filter(abs(rt-initial_RT) <= initial_tolerance) |>
    filter(abs(mz-target_MZ) <= tolerance) |>
    group_by(rt) |>
    summarise(int = max(int)) 
  
  subset_data <- data |>
    filter(abs(rt-initial_RT) <= 2/100 * initial_RT) 
  
  subset_min <- min(subset_data$rt)
  subset_max <- max(subset_data$rt)
  
  final_RT <- subset_data |>
    arrange(desc(int)) |>
    head(1) |>
    pull(rt)
  
  int_target <- subset_data |>
    arrange(desc(int)) |>
    head(1) |>
    pull(int)
  
  final_tolerance <- 1/100 * final_RT
  
  final_data <- data |>
    filter(abs(rt-final_RT) <= final_tolerance)
  
  min_bound <- data |>
    filter(rt <= final_RT) |>
    arrange(desc(rt)) |> 
    filter(lead(int) < int_target*0.2) |> 
    filter(int < int_target*0.2) |>
    head(1) |>
    pull(rt)
  
  max_bound <- data |>
    filter(rt >= final_RT) |>
    arrange(rt) |> 
    filter(lead(int) < int_target*0.2) |> 
    filter(int < int_target*0.2) |> 
    head(1) |>
    pull(rt)
  

  p <- ggplot(final_data, aes(x = rt, y = int)) +
        geom_line(color = inq_dark_blue, show.legend = FALSE) + 
        geom_ribbon(data = subset(data, rt >= min_bound & rt <= max_bound),
                    aes(ymin = 0, ymax = int, color = "Quantification area"),
                    alpha = 0.3, fill = inq_blue, linewidth = 0) +
        geom_vline(aes(xintercept = final_RT, color = "Final identification"), 
                   linetype = "solid", size = 0.8) + 
        geom_vline(aes(xintercept = min_bound, color = "Boundary"), 
                   linetype = "dashed", size = 0.8) +
        geom_vline(aes(xintercept = max_bound, color = "Boundary"), 
                   linetype = "dashed", size = 0.8) +
        geom_hline(aes(yintercept = int_target * 0.2, color = "20% intensity threshold"), 
                   linetype = "dotted", size = 1) +
        scale_color_manual(
          name = NULL,
          values = c(
            "Final identification" = inq_pink,
            "Boundary" = inq_green,
            "20% intensity threshold" = inq_purple,
            "Quantification area" = inq_blue
          )
        ) +
        labs(
          subtitle = str_c('File: ',
                           fileID,
                           ' | Final RT: ', 
                           round(final_RT, digits = 2), 
                           ' min',
                           ' | Final RT window: +/-',
                           round(final_RT*0.01*60, digits = 2),
                           ' seconds'),
          x = 'RT [min]',
          y = 'Intensity'
        ) +
        theme_minimal() +
        theme(
          legend.position = c(0.98, 0.98),  
          legend.justification = c("right", "top"),
          legend.background = element_rect(fill = "white", color = "white"),
          legend.box.background = element_rect(color = "white"),
          legend.key.height = unit(0.7, "cm"),
          legend.spacing.y = unit(-0.2, "cm"),
          plot.background = element_rect(color = 'white')
        ) +
        guides(
          color = guide_legend(
            override.aes = list(
              size = 3  
            )
          )
        )
  
  return(p)
  
}
```

### Plot MBR Bar plot

```{r}

plot_bar_MBR_count <- function(psms, name) {
  
p <- psms |>
  mutate(MBR = ifelse(is.na(conf), 'MBR',
                      ifelse(is.na(abundance), 'Not quantified', 'Quantified'))) |>
  count(file_id, MBR) |>
  group_by(file_id) |>
  mutate(percentage = (n / sum(n)) * 100) |>
  mutate(MBR = factor(MBR,
                      levels = MBR |>
                        table() |>
                        sort(decreasing = FALSE) |>
                        names())) |>
  ggplot(aes(x = file_id, y = n, fill = MBR)) +
  geom_bar(stat = 'identity', alpha = 0.8) +
  scale_fill_manual(values = c('MBR' = inq_purple,
                               'Quantified' = inq_green,
                               'Not quantified' = 'grey')) +
  geom_label(aes(label = str_c(round(percentage, digits = 1), '%')),
             position = position_stack(vjust = 0.5),
             color = 'white',
             size = 3,
             alpha = 0.6,
             label.size = 0,
             show.legend = FALSE) +
  theme_minimal() +
  labs(title = str_c(name, ': Identified predictions by InstaNovo and MBR'),
       x = 'Experiment',
       y = 'Count',
       fill = '')

  return(p)
}

```

### Log2 fold change plot for the mixed proteome experiment

```{r}

plot_fold_change_mixed_proteome <- function(table, 
                                      ylims = c(-8,8), 
                                      xlims = c(10,40),
                                      abundance_pattern,
                                      mbr, 
                                      type, 
                                      titles = TRUE,
                                      custom_titles = '') {
  
  if (type == 'Protein'){
  fc_data <- table |>
    group_by(principal_protein) |>
    mutate(protein_group = str_c(unique(protein_group), collapse = ';')) |>
    ungroup() |>
    distinct() |>
    select(principal_protein, species, contains('ormalized')) |>
    pivot_longer(cols = contains('ormalized'),
                 names_to = c("experiment"),
                 names_pattern = abundance_pattern,
                 values_to = "abundance") |>
    filter(!str_detect(experiment, '1')) |>
    mutate(abundance = ifelse(abundance == 0, NA, abundance)) |>
    mutate(experiment = str_extract(experiment,'[AB]')) |>
    group_by(principal_protein, species, experiment) |>
    summarise(abundance_comb = mean(abundance, na.rm = TRUE), .groups = 'drop') |>
    mutate(abundance_comb = na_if(abundance_comb, NaN)) |>
    pivot_wider(names_from = experiment,
                values_from = abundance_comb) |>
    filter(!if_all(c(A,B), is.na)) |>
    drop_na() |>
    mutate(fold_change = A/B,
           mean_abundance = (A+B)/2)
  } else {
    fc_data <- table |>
    select(sequence, meas_mass, species, contains('ormalized') ) |>
    pivot_longer(cols = contains('ormalized'),
                 names_to = c("experiment"),
                 names_pattern = abundance_pattern,
                 values_to = "abundance") |>
    mutate(abundance = ifelse(abundance == 0, NA, abundance)) |>
    drop_na(species) |>
    drop_na(abundance) |>
    filter(!str_detect(experiment, '1')) |>
    mutate(experiment = str_extract(experiment,'[AB]')) |>
    group_by(sequence, species, experiment, meas_mass) |>
    summarise(abundance_comb = mean(abundance, na.rm = TRUE), .groups = 'drop') |>
    pivot_wider(names_from = experiment,
                values_from = abundance_comb) |>
    drop_na() |>
    mutate(A = replace_na(A,1),
           B = replace_na(B,1)) |>
    mutate(fold_change = A/B,
           mean_abundance = (A+B)/2)
  }
    
   if (custom_titles != ''){
     title_labs <- labs(title = custom_titles,
                        color = 'Species')
   }
   else if (titles) {
    title_labs <- labs(title = str_c('Mixed proteome: ', type,
                                     ' fold change over mean normalized abundance'),
         subtitle = str_c('for mean(A2,A3)/mean(B2,B3) with NAs dropped', mbr),
         color = 'Species') 
    } else {
      title_labs <- labs(color = 'Species')
    }
  # Main scatterplot
  scatter_plot <- fc_data |>
    ggplot(aes(x = log2(mean_abundance), 
               y = log2(fold_change), 
               color = species)) +
    geom_point() + 
    geom_hline(yintercept = c(2, 0, -1), 
               linetype = 'dashed', 
               color = c(inq_pink, inq_orange, inq_purple)) +
    title_labs +
    scale_color_manual(values = c(ECOLI = inq_pink,
                                  HUMAN = inq_orange,
                                  YEAST = inq_purple),
                       guide = guide_legend(override.aes = list(shape = 15))) + 
    coord_cartesian(ylim = ylims, xlim = xlims) +
    theme_minimal() +
    theme(legend.position = 'none')
  
  # Boxplot of log2(fold_change) per species
  box_plot <- fc_data |>
    ggplot(aes(x = species, y = log2(fold_change), color = species)) +
    geom_boxplot(width = 0.6, show.legend = FALSE) +
    geom_point(aes(x = species, y = log2(fold_change), color = species), 
               alpha = 0, 
               show.legend = TRUE) +
    geom_hline(yintercept = c(2, 0, -1), 
               linetype = 'dashed', 
               color = c(inq_pink, inq_orange, inq_purple)) +
    scale_color_manual(values = c(ECOLI = inq_pink,
                                  HUMAN = inq_orange,
                                  YEAST = inq_purple),
                       guide = guide_legend(override.aes = list(alpha = 1, 
                                                                shape = 15,
                                                                size = 5))) +
    coord_cartesian(ylim = ylims) +
    theme_void() +
    labs(color = 'Species') +
    theme(
      axis.title = element_blank(),
      axis.text = element_blank(),
      axis.ticks = element_blank(),
      legend.position = "right"
    ) 
  
  p <- scatter_plot + box_plot + plot_layout(widths = c(3, 1), axis_titles = 'collect')
  
  return(p)

}
```

### Venn diagram

```{r}

plot_venn_diagram <- function(name1, name2, list1, list2) {
  fill_colors <- c(inq_pink, inq_blue)
 
  for_venn <- setNames(list(list1, list2), c(name1, name2))
  
  venn <- ggvenn(
    for_venn,
    fill_color = fill_colors,
    stroke_size = 0,
    set_name_size = 0,
    auto_scale = TRUE
  )
  
  names(fill_colors) <- c(name1, name2)
  legend_data <- data.frame(group = factor(names(fill_colors), levels = names(fill_colors)))

  legend_plot <- ggplot(legend_data, aes(x = 0.01, y = group, fill = group)) +
    geom_tile(width = 0, height = 0, alpha = 0.5) +
    scale_fill_manual(name = NULL, values = fill_colors) +
    theme_void() +
    theme(
      legend.position = "bottom",
      legend.text = element_text(size = 13),
      plot.margin = margin(10, 0, 10, 0)
    )
  
  combined_plot <- venn / legend_plot + plot_layout(heights = c(4,0))
  
  return(combined_plot)
}
```

### Calibration scatter plot trends

```{r}
plot_density_point <- function(df, x, title, ylim) {
  df <- df |> 
    filter(!is.na(ppm_iso), 
           !is.na(ppm_cali))
  
  if(x == "rt"){
    df <- df |> 
      mutate(x = rt)
    xlab <- "Retention time (s)"
  }
  else{
    df <- df |> 
      mutate(x = meas_mz)
    xlab <- "Measured m/z"
  }
  
  if(title == "Calibrated"){
    df <- df |> 
      mutate(ppm = ppm_cali)
  }
  else{
    df <- df |> 
      mutate(ppm = ppm_iso)
  }
  
  y_min_ppm = floor(min(df$ppm_iso, df$ppm_cali))
  y_max_ppm = ceiling(max(df$ppm_iso, df$ppm_cali))

  
  plot <- df |> 
    ggplot(aes(x = x, 
               y = ppm)) +
    geom_pointdensity(size = 0.5) +
    scale_color_gradient(low = inq_orange, high = inq_purple) +
    geom_hline(aes(yintercept = 0), size = 0.5, linetype = "dashed", color = "black") + 
    geom_smooth(method = "loess", se = FALSE, color = inq_dark_blue, size = 0.7) +
    coord_cartesian(ylim = c(-1*ylim, ylim)) +
    theme_minimal() +
    theme(legend.position = "none") + 
    labs(x = xlab, y = "Error (ppm)", title = title)
  
  return(plot)
}

```

### Calibration boxplot

```{r}
plot_boxplot_calibration <- function(df, outliers = FALSE) {
  
  df |> 
  filter(!is.na(correct)) |>
    pivot_longer(cols = c(iso_adj_mz_error, calibrated_mz_error), 
                   names_to = "type", values_to = "value") |> 
    mutate(type = factor(type,
                         levels = c("iso_adj_mz_error", 
                                    "calibrated_mz_error"))) |>  
    ggplot(aes(x = type, y = value, fill = correct)) + 
    geom_boxplot(alpha = 0.7, outliers = outliers) + 
    scale_fill_manual(values = c("TRUE" = inq_green,
                                 "FALSE" = inq_blue), 
                      labels = c("TRUE" = "Correct",
                                 "FALSE" = "Wrong")) +
    labs(x = "", 
         y = "m/z error", 
         fill = "Identification",
         title = "Error in m/z before and after calibration") +
    scale_x_discrete(
      labels = c("iso_adj_mz_error" = "After isotope\nadjustment",
                 "calibrated_mz_error" = "After calibration")) +
    theme_minimal() +
  theme(title = element_text(size = 8.5),
        legend.title = element_text(size = 8.5),
        legend.text = element_text(size = 7.5),
        legend.position = "bottom")
}
```

### Calibration density

```{r}

plot_density_calibration <- function(df, calibrated = TRUE) {
  
  if(calibrated){
    df <- df |> 
      mutate(mz_error = calibrated_mz_error)
    title = "After calibration"
  }
  else{
    df <- df |> 
      mutate(mz_error = iso_adj_mz_error)
    title = "Before calibration"
  }
  
  df |> 
    filter(!is.na(correct)) |>
    ggplot() +
    geom_vline(xintercept = 0, linetype = "dashed", color = "red", size = 0.6) +
    geom_density(aes(mz_error,  
                       color = correct), size = 0.7) + 
      scale_color_manual(values = c("TRUE" = inq_green,
                                   "FALSE" = inq_blue), 
                        labels = c("TRUE" = "Correct",
                                   "FALSE" = "Wrong")) +
    coord_cartesian(xlim = c(-0.01, 0.01)) +
    labs(
      title = title, 
      x = "mz error",
      y = "Density",
      color = "Identification"
    ) +
    theme_minimal() +
    theme(
      axis.title = element_text(size = 12),
      legend.title = element_text(size = 12),
      legend.text = element_text(size = 10)
    )
}
```

### C-terminal amino acid distribution plot

```{r}

plot_c_term_AA_distribution <- function(peptides, title, abundance_pattern) {
  
  p <- peptides |>
    select(sequence, contains('ormalized')) |>
    pivot_longer(cols = contains('ormalized'),
                   names_to = c("experiment"),
                   names_pattern = abundance_pattern,
                   values_to = "abundance") |>
    drop_na(abundance) |>
    mutate(experiment = str_extract(experiment, '\\D+'),
           c_term_cleav = str_detect(sequence, '[KRED]$')) |>
    mutate(c_term_cleav = ifelse(c_term_cleav, str_extract(sequence, '\\D{1}$'),'Other')) |>
    mutate(c_term_cleav = factor(c_term_cleav,
                        levels = c('E','D','K','R','Other'))) |>
    group_by(experiment, c_term_cleav) |>
    summarise(abundance_comb = sum(abundance, na.rm = TRUE), .groups = 'keep') |>
    mutate(abundance_comb = na_if(abundance_comb, 0)) |>
    summarise(total_abundance = sum(abundance_comb, na.rm = TRUE), .groups = 'drop') |>
    group_by(experiment) |>
    mutate(total_abundance_experiment = sum(total_abundance)) |>
    ungroup() |>
    mutate(abundance_pct = total_abundance / total_abundance_experiment * 100) |>
    mutate(label = str_c(round(abundance_pct, digits = 1), '%')) |>
    select(-total_abundance_experiment) |>
    ggplot(aes(x = experiment, y = abundance_pct, fill = c_term_cleav)) +
    geom_col(width = 0.6, alpha = 0.8) +
    geom_label(aes(label = ifelse(abundance_pct > 0.2, label, NA)),
              position = position_stack(vjust = 0.5),
              size = 3, 
              color = "white",
              label.size = 0,
              alpha = 0.6,
              show.legend = FALSE,
              na.rm = TRUE) +
    scale_fill_manual(values = c('E' = inq_green,
                                 'D' = inq_purple,
                                 'K' = inq_blue,
                                 'R' = inq_dark_blue,
                                 'Other' = 'grey')) + 
    
    theme_minimal() + 
    theme(legend.position = 'right',
          legend.title.position = 'top',
          legend.direction = 'vertical') +
    labs(title = title,
         x = 'Experiment',
         y = 'Abundance [%]',
         fill = 'C-Terminal \nAmino Acid') 
  
  return(p)
}
```

### Species distribution

```{r}

plot_species_distribution <- function(peptides, 
                                      title,
                                      subtitle = '', 
                                      abundance_pattern) {
  
  if (subtitle != ''){
  custom_labs <- 
    labs(title = title,
         subtitle = subtitle,
         x = 'Experiment',
         y = 'Abundance [%]',
         fill = 'Species')
  } else {
    custom_labs <- labs(title = title,
         x = 'Experiment',
         y = 'Abundance [%]',
         fill = 'Species') 
  }

  p <- peptides |>
    select(sequence, meas_mass, species, contains('ormalized')) |>
    pivot_longer(cols = contains('ormalized'),
                   names_to = c("experiment"),
                   names_pattern = abundance_pattern,
                   values_to = "abundance") |>
    mutate(abundance = ifelse(abundance == 0, NA, abundance)) |>
    drop_na(abundance) |>
    filter(!str_detect(experiment, '1')) |>
    mutate(experiment = str_extract(experiment,'[AB]')) |>
    group_by(sequence, species, experiment, meas_mass) |>
    summarise(abundance_comb = sum(abundance, na.rm = TRUE), .groups = 'drop') |>
    mutate(abundance_comb = na_if(abundance_comb, 0)) |>
    group_by(experiment, species) |>
    summarise(total_abundance = sum(abundance_comb, na.rm = TRUE), .groups = 'drop') |>
    group_by(experiment) |>
    mutate(total_abundance_experiment = sum(total_abundance)) |>
    ungroup() |>
    mutate(abundance_pct = (total_abundance / total_abundance_experiment) * 100) |>
    mutate(label = str_c(round(abundance_pct, digits = 1),'%')) |>
    select(-total_abundance_experiment) |>
    ggplot(aes(x = experiment, y = abundance_pct, fill = species)) +
    geom_col(alpha = 0.9) +
    geom_label(aes(label = ifelse(abundance_pct > 1, label, NA)),
              position = position_stack(vjust = 0.5),
              size = 3, 
              color = "white",
              label.size = 0,
              alpha = 0.6,
              show.legend = FALSE,
              na.rm = TRUE) +
    scale_fill_manual(values = c('HUMAN' = inq_orange,
                                 'YEAST' = inq_purple,
                                 'ECOLI' = inq_pink,
                                 'NA' = 'grey')) + 
    custom_labs +
    theme_minimal() + 
    theme(legend.position = 'right',
          legend.title.position = 'top',
          legend.direction = 'vertical')
  
  return(p)
}
```

### Protein Groups

```{r}

plot_groups <- function(ungrouped_protein_table,
                        protein_group_table,
                        selected_protein,
                        show_protein_names = FALSE) {
  selected_protein <- fixed(selected_protein)

plotting_table <- ungrouped_protein_table |>
    mutate(protein_type = ifelse(str_detect(accession,'^B_fig'),
                               'Native', 
                               ifelse(str_detect(accession, '^P'), 
                                      'Phage', 'Anti-viral'))) |> 
    select(accession, no_peptides, no_unique_peptides_protein, peptide_seqs, protein_type) |>
    mutate(peptide_seqs = str_split(peptide_seqs, ";")) |>
    unnest(peptide_seqs)
  
  # Initialize protein group based on selected protein
  protein_group <- protein_group_table |>
      filter(str_detect(protein_group, selected_protein) |
           str_detect(principal_protein, selected_protein))
  
  repeat {
      # Extract peptides from the current protein group
      peptide_group <- protein_group |>
          select(peptide_seqs) |>
          mutate(peptide_seqs = str_split(peptide_seqs, ";")) |>
          unnest(peptide_seqs) |> 
          pull(peptide_seqs) |> 
          unique()  # Get unique peptides as a vector
      
      found_proteins <- protein_group |>
        select(protein_group, principal_protein) |> 
        mutate(Proteins = ifelse(is.na(protein_group), principal_protein,
                                 str_split(protein_group, ';'))) |>
        select(Proteins) |>
        unnest(Proteins) |>
        pull(Proteins) |>
        unique()
      
      # Remove empty strings from found protein
      found_proteins <- found_proteins[found_proteins != ""]
        
      # Find all proteins connected to these peptides
  
      new_protein_group <- protein_group_table |> 
        rowwise() |> 
        filter(
          any(str_detect(peptide_seqs, peptide_group)) |
          any(str_detect(fixed(principal_protein), fixed(found_proteins))) |
          any(str_detect(fixed(protein_group), fixed(found_proteins)))) |> 
        ungroup()

      
      # Stop if no new proteins were found
      if (nrow(new_protein_group) == nrow(protein_group)) {
        break
      }
      # Expand protein group
      protein_group <- new_protein_group
  }
  
  # Extract Master proteins
  protein_masters <- protein_group |> 
    select(principal_protein) |> 
    distinct() |> 
    pull()
  
  # Extract all individual proteins from the protein group
   protein_group <- protein_group |>
      select(protein_group, principal_protein) |> 
      mutate(Proteins = ifelse(is.na(protein_group), principal_protein,
                               str_split(protein_group, ';'))) |>
      select(Proteins) |>
      unnest(Proteins)
    protein_group <- protein_group$Proteins
    
    protein_group <- protein_group[protein_group != ""]

  
  # Create groupings for convex hulls
  groupings <- list()
  
  for (protein in protein_masters) {
    this_group <- protein_group_table |>
       filter(str_detect(principal_protein, protein)) |>
       select(principal_protein, protein_group, peptide_seqs) |>
       mutate(peptide_seqs = str_split(peptide_seqs, ";"),
              Proteins = str_split(protein_group, ';'),
              Master_protein = principal_protein) |>
       select(Master_protein, peptide_seqs, Proteins) |>
       unnest(peptide_seqs) |>
       unnest(Proteins) |>
       pivot_longer(c(Master_protein, Proteins, peptide_seqs), names_to = 'All') |>
       pull(value) |>
       unique()
    groupings[[protein]] <- this_group
  }
  
  # Convert data into an edge list (Protein - Peptide relationships)
  edge_list <- plotting_table |>
    select(accession, peptide_seqs) |>
    unnest(peptide_seqs)
  
  graph <- graph_from_data_frame(edge_list, directed = FALSE)
  
  # Identify connected components (clusters)
  clusters <- components(graph)
  
  # Add cluster information as a vertex attribute
  V(graph)$cluster <- clusters$membership
  
  # Identify node types (Protein or Peptide)
  V(graph)$type <- ifelse(V(graph)$name %in% protein_masters, 
                          "Master Protein", 
                          ifelse(V(graph)$name %in% protein_group,
                                 'Protein', "Peptide"))
  
  # Ensure correct Unique Peptides mapping using left_join
  unique_peptide_df <- plotting_table |>
    select(accession, no_unique_peptides_protein) |>
    distinct()
  
  unique_peptide_map <- deframe(unique_peptide_df)
  
  V(graph)$no_unique_peptides_protein <- ifelse(
    (V(graph)$type == "Protein" | V(graph)$type == "Master Protein"),
    unique_peptide_map[V(graph)$name],  # Lookup correct value
    NA  # Peptides get NA
  )
  
  # Convert cluster membership into a dataframe
  cluster_df <- data.frame(
    name = V(graph)$name,
    cluster = clusters$membership
  )
  
  # Find the cluster of the selected protein
  selected_cluster <- filter(cluster_df, name == selected_protein)$cluster

  # Get all the nodes connected to the selected protein's cluster
  connected_nodes <- neighborhood(graph, order = 1, nodes = protein_group)[[1]]
  
  # Get clusters of connected nodes
  connected_clusters <- unique(clusters$membership[connected_nodes])
  
  # Filter graph to include all nodes from connected clusters
  subgraph_nodes <- cluster_df |>
    filter(cluster %in% connected_clusters) |>
    pull(name)
  
  subgraph <- induced_subgraph(graph, vids = subgraph_nodes)
  
  # Ensure correct Protein/Peptide classification
  V(subgraph)$type <- ifelse(V(subgraph)$name %in% protein_masters, 
                          "Master Protein", 
                          ifelse(V(subgraph)$name %in% protein_group,
                                 'Protein', "Peptide"))
  
  lookup_table <- plotting_table |> 
  filter(accession %in% V(subgraph)$name) |> 
  select(accession, protein_type) |> 
  unique()
  
  V(subgraph)$color_label <- factor(ifelse(
    (V(subgraph)$type == "Protein" | V(subgraph)$type == "Master Protein"),
    lookup_table$protein_type[match(V(subgraph)$name, lookup_table$accession)],
  "Peptide"
))

  # Define colors manually for legend'
  color_mapping  <- c('Anti-viral' = inq_purple,
                      'Native' = inq_dark_blue,
                      'Phage' = inq_orange)
  
  type_mapping <- c('Master Protein' = 15,
                    'Protein' = 17,
                    'Peptide' = 16)
  V(subgraph)$label <- ifelse(
    (V(subgraph)$type == "Protein" | 
       V(subgraph)$type == "Master Protein") & show_protein_names, 
    V(subgraph)$name, 
    ""
    # V(subgraph)$name
  )
  
  # Extract graph layout
  graph_layout <- create_layout(subgraph, layout = "stress")

  # Extract coordinates from the layout
  layout_df <- as.data.frame(graph_layout)

  # Compute convex hulls for each group
  hull_df <- bind_rows(
    lapply(names(groupings), function(group) {
      group_nodes <- groupings[[group]] 
      group_layout <- layout_df |>
        filter(name %in% group_nodes) |>
        select(x, y, name) |>
        mutate(Group = group)
      
      # Compute convex hull if at least 3 points exist
      if (nrow(group_layout) >= 3) {
        hull_points <- chull(group_layout$x, group_layout$y)
        expanded_hull <- group_layout[hull_points, ]
        
        # Apply padding by shifting points outward
        expanded_hull <- expanded_hull |>
          mutate(x = x + 0.07 * (x - mean(x)),  # Expand outward
                 y = y + 0.07 * (y - mean(y)))  # Expand outward
        
        expanded_hull
      } else {
        group_layout
      }
    })
  )

  # Ensure correct protein names
  protein_name_map <- unique_peptide_df |> deframe()
  
  # Plot with convex hulls
  ggraph(graph_layout) +  
    geom_polygon(data = hull_df, aes(x = x, y = y, group = Group),
                 alpha = 0, color = "black") +  
    geom_edge_link(color = "gray80", alpha = 0.5) +  
    geom_node_point(aes(color = color_label, shape = type, size = type)) + 
    geom_node_text(aes(label = label), size = 4, vjust = -0.5, repel = TRUE) +  
    scale_shape_manual(values = type_mapping, name = "Type") +  
    scale_color_manual(values = color_mapping, name = "Protein species") +
    scale_size_manual(values = c(4,1,4), guide = "none") +
    labs(title = str_c('Protein group containing ', selected_protein)) + 
    theme(legend.position = "bottom", legend.direction = "vertical") + 
    theme(panel.background = element_rect(fill = "white", color = "white"))
  
}

```

## Plot display

### The mixed proteome experiment

#### Quantification spectras with MBR

```{r}

seq <- 'GACLLPK'

# Extract the data for the plots
assign(str_c(seq, '_data'), mixed_MBR_preds |>
  filter(preds == seq) |>
  select(preds, file_ID, mz, rt, log_probs)
)

order <- get(str_c(seq,'_data')) |>
  arrange(is.na(log_probs)) |>
  pull(file_ID)

assign(str_c(seq,'_comparison'), 
       plot_MBR_comparison(seq, 
        next_group = 1, 
        exp_order = order)
)

save_fig(get(str_c(seq,'_comparison')), 'GACLLPK_combined_MBR.png', h = 1440*1.4)

```

```{r, warning=FALSE}

for (i in 1:nrow(get(str_c(seq,'_data')))) {
  
  row <- get(str_c(seq,'_data'))[i,]
  file_id <- row$file_ID
  MBR <- is.na(row$log_probs)
  
  if (!MBR) {
    assign(str_c(seq,'_',file_id), 
           plot_quantification_area(
                specs = get(str_c('mixed_specs_',file_id)),
                 preds = mixed_preds,
                 target_seq_spec = seq,
                 fileID = file_id
                )
           )
  } else {
    
    mz <- row$mz
    rt <- row$rt
    
    assign(str_c(seq,'_',file_id), 
           plot_quantification_area_MBR(
                specs = get(str_c('mixed_specs_',file_id)),
                 seq = seq,
                 fileID = file_id,
                 target_MZ = mz,
                 initial_RT = rt
                )
           )
    
    assign(str_c(seq,'_',file_id,'_MBR'), 
           plot_quantification_area_MBR_full(
                specs = get(str_c('mixed_specs_',file_id)),
                 seq = seq,
                 fileID = file_id,
                 target_MZ = mz,
                 initial_RT = rt
                )
           )
  }
}
```

Combining the plots for comparison:

```{r, output = FALSE}

MBR_false <- get(str_c(seq,'_data')) |>
  filter(!is.na(log_probs)) |>
  pull(file_ID)

MBR_true <- get(str_c(seq,'_data')) |>
    filter(is.na(log_probs)) |>
    pull(file_ID)

assign(str_c(seq,'_identified'), Reduce(`+`,mget(str_c(seq,'_',MBR_false))) + 
  plot_layout(1,3, guides = 'collect', 
              axis_titles = 'collect') +
  plot_annotation(title = 'Mixed proteome: InstaNovo predictions for GACLLPK') & 
  theme(legend.position = 'bottom')
)

assign(str_c(seq,'_MBR'), Reduce(`+`,mget(str_c(seq,'_',MBR_true,'_MBR'))) +
  plot_layout(1,3, guides = 'collect', axis_titles = 'collect') +
  plot_annotation(title = 'Mixed proteome: MBR predictions for GACLLPK with initial RT') & 
  theme(legend.position = 'bottom')
)

assign(str_c(seq,'_MBR_final'), Reduce(`+`,mget(str_c(seq,'_',MBR_true))) + 
  plot_layout(1,3, guides = 'collect', axis_titles = 'collect') +
  plot_annotation(title = 'Mixed proteome: MBR predictions for GACLLPK') & 
  theme(legend.position = 'bottom') 
)

```

```{r}

save_fig(get(str_c(seq,'_identified')), 
         str_c(seq,'_identified.png'), 
         w = 1920*1.1)
save_fig(get(str_c(seq,'_MBR')), 
         str_c(seq,'_MBR.png'), 
         w = 1920*1.2, 
         h = 1440*1.1)
save_fig(get(str_c(seq,'_MBR_final')), 
         str_c(seq,'_MBR_final.png'), 
         w = 1920*1.1)

```

#### Comparing MBR and non-MBR abundances

```{r}

save_fig(plot_bar_MBR_count(mixed_MBR_psms, 'Mixed proteome'), 
              'mixed_MBR_bar_plot.png', 
              h = 1340)

```

#### INQuant and database search log2 fold change for different species with boxplots

```{r}

custom_xlims <- c(14,32)

fold_change_peptide_inquant <- 
  plot_fold_change_mixed_proteome(
    table = mixed_peptide |> 
      mutate(species = mixed_fasta$species
          [match(str_split_i(principal_protein,';',1), 
                 mixed_fasta$accession)]), 
    abundance_pattern = 'abundance_(.*)_normalized', 
    type = 'Peptide', 
    custom_title = 'INQuant', 
    xlims = custom_xlims) 

fold_change_peptide_database <- plot_fold_change_mixed_proteome(
  table = mixed_database_peptides |> 
    mutate(species = mixed_fasta$species
          [match(str_split_i(`Master Protein Accessions`,';',1), 
                 mixed_fasta$accession)],
          sequence = `Annotated Sequence`,
          meas_mass = `Theo. MH+ [Da]`), 
  abundance_pattern = 'Abundances \\(Normalized\\): (.*): Sample, [AB]', 
  type = 'Peptide', custom_title = 'Database', 
  xlims = custom_xlims)


mixed_fold_change_peptide_combined <- 
  ((fold_change_peptide_inquant / fold_change_peptide_database) +
     plot_layout(1, 2, axis_titles = 'collect', guides = 'collect') + 
     plot_annotation(title = 
          'Mixed proteome: Peptide fold change over mean normalized abundance',
        subtitle = 'for mean(A2,A3)/mean(B2,B3) with NAs dropped'))

save_fig(mixed_fold_change_peptide_combined, 
              'mixed_fold_change_plot_inquant_vs_database_combined.png', 
              h = 2300)
```

```{r}

psms_only_MBR <- mixed_MBR_psms |>
  filter(is.na(conf)) |>
  pull(sequence) |>
  unique()

mixed_only_MBR_peptides <- mixed_MBR_peptide |>
 mutate(species = mixed_fasta$species
          [match(str_split_i(principal_protein,';',1), 
                 mixed_fasta$accession)]) |>
  filter(sequence %in% psms_only_MBR)
  
mixed_peptide_fold_change_only_MBR <- 
  plot_fold_change_mixed_proteome(table = mixed_only_MBR_peptides,
                            custom_title = 'Only MBR',
                            abundance_pattern = 'abundance_(.*)_normalized',
                            type = 'Peptide',
                            xlims = custom_xlims)

mixed_peptide_fold_change_MBR <- 
  plot_fold_change_mixed_proteome(table = mixed_MBR_peptide |>
                                     mutate(species = mixed_fasta$species
                                        [match(str_split_i(principal_protein,';',1), 
                                          mixed_fasta$accession)]),
                            custom_title = 'With MBR',
                            abundance_pattern = 'abundance_(.*)_normalized',
                            type = 'Peptide',
                            xlims = custom_xlims)

mixed_peptide_fold_change_no_MBR <- 
  plot_fold_change_mixed_proteome(table = mixed_peptide |>
                                     mutate(species = mixed_fasta$species
                                        [match(str_split_i(principal_protein,';',1), 
                                          mixed_fasta$accession)]),
                            custom_title = 'Without MBR',
                            abundance_pattern = 'abundance_(.*)_normalized',
                            type = 'Peptide',
                            xlims = custom_xlims)

mixed_peptide_fold_change_MBR_combined <- 
  (
    (mixed_peptide_fold_change_no_MBR / 
      mixed_peptide_fold_change_MBR  / 
      mixed_peptide_fold_change_only_MBR) +
  plot_layout(1, 3, axis_titles = 'collect', guides = 'collect') + 
  plot_annotation(
    title = 
        'Mixed proteome: Peptide fold change over mean normalized abundance',
    subtitle = 
        'for mean(A2,A3)/mean(B2,B3) with NAs dropped')
  )

save_fig(mixed_peptide_fold_change_MBR_combined, 
              'mixed_peptide_fold_change_combined.png', 
              h = 1440*2)
```

#### Peptide and protein log2 fold change

```{r, output = FALSE}

fold_change_peptide_inquant_empty_MBR <- 
  plot_fold_change_mixed_proteome(table = mixed_MBR_peptide |> 
    mutate(species = mixed_fasta$species
          [match(str_split_i(principal_protein,';',1), 
                 mixed_fasta$accession)]), 
          abundance_pattern = 'abundance_(.*)_normalized', 
          type = 'Peptide', 
          custom_title = 'Peptides', 
          xlims = custom_xlims) 

fold_change_protein_inquant_empty_MBR <- 
  plot_fold_change_mixed_proteome(table = mixed_MBR_proteins |> 
                            mutate(species = mixed_fasta$species
                                  [match(principal_protein, 
                                         mixed_fasta$accession)]),
          abundance_pattern = 'abundance_(.*)_normalized', 
          type = 'Protein', 
          custom_title = 'Proteins', 
          xlims = custom_xlims) 

mixed_fold_change_combined_protein_peptide <- ((fold_change_peptide_inquant_empty_MBR  / fold_change_protein_inquant_empty_MBR) +
plot_layout(1, 2, axis_titles = 'collect', guides = 'collect') + 
plot_annotation(title = 
          'Mixed proteome: Fold change over mean normalized abundance',
        subtitle = 'for INQuant (with MBR) mean(A2,A3)/mean(B2,B3) with NAs dropped'))

```

```{r}

save_fig(mixed_fold_change_combined_protein_peptide, 
              'mixed_fold_change_plot_peptide_vs_protein_combined.png', 
              h = 2500)

```

#### Abundance distribution for each species for INQuant and database search

```{r, output = FALSE}

mixed_species_distribution_no_NA <- 
  plot_species_distribution(
    mixed_peptide |>
     mutate(species = mixed_fasta$species
            [match(str_split_i(principal_protein,';',1), 
                   mixed_fasta$accession)]) |>
      drop_na(species), 
    title = 'INQuant', 
    subtitle = '(only peptides \nmapping to protein)',
    abundance_pattern = 'abundance_(.*)_normalized') + 
  theme(legend.position = 'none')

mixed_species_distribution <- 
  plot_species_distribution(
    mixed_peptide |>
      mutate(species = mixed_fasta$species
          [match(str_split_i(principal_protein,';',1), 
                 mixed_fasta$accession)]), 
   title = 'INQuant', 
   abundance_pattern = 'abundance_(.*)_normalized') 

mixed_database_species_distribution <- 
  plot_species_distribution(
    mixed_database_peptides |>
      mutate(sequence = `Annotated Sequence`, 
             meas_mass = `Theo. MH+ [Da]`), 
  title = 'Database', 
  abundance_pattern = 
    'Abundances \\(Normalized\\): (.*): Sample, [AB]') 

mixed_species_distribution_combined <- 
  (mixed_species_distribution + 
     mixed_species_distribution_no_NA + 
     mixed_database_species_distribution +
  plot_layout(3, 1, 
              axis_titles = 'collect', 
              guides = 'collect') + 
  plot_annotation(
    title = 
      'Mixed proteome: Peptide abundance percentage distribution',
    subtitle = 
      'For each experiment, categorised by species, not including A1'
    )
  )

```

```{r}

save_fig(mixed_species_distribution_combined, 
              'mixed_database_inquant_species_distribution.png',
              h = 1600)

```

#### Identified peptides Venn diagram

Sequences get added ID's to ensure that identical sequences, which are not identical in m/z are counted separately in the Venn diagram.

```{r, output = FALSE}

mixed_identified <- mixed_MBR_peptide |>
  filter(!if_all(contains('abundance'), ~ is.na(.))) |>
  mutate(sequence = str_c(sequence, "_", row_number()), .by = sequence) |>
  pull(sequence)

mixed_database_identified <- mixed_database_peptides |>
  filter(!if_all(contains('Abundance'), ~ is.na(.))) |>
  mutate(Sequence = str_split_i(`Annotated Sequence`,'\\.',2)) |>
  mutate(Sequence = str_replace_all(Sequence, 'I','L')) |>
  mutate(sequence = str_c(Sequence, "_", row_number()), .by = Sequence) |>
  pull(sequence)

mixed_venn <- plot_venn_diagram('INQuant', 'Database',
                  mixed_identified,
                  mixed_database_identified) +
  plot_annotation(title = 'Quantified peptides in the mixed proteome experiment',
                  subtitle = 'With Database Is replaced by Ls',
                  theme = theme(plot.title = element_text(hjust = 0.5),
                                plot.subtitle = element_text(hjust = 0.5)))

```

```{r}

save_fig(mixed_venn, 'mixed_venn_identified_peptides.png', 
              w = 1600,
              h = 1600)

```

#### Average amount of quantification values per peptide

```{r}

mixed_identified_per_peptide <- 
  mixed_MBR_peptide |>
  filter(!if_all(contains('abundance'), ~ is.na(.))) |>
  select(sequence, contains('abundance') & !contains('normalized')) |>
  rowwise() |>
  mutate(amount_quant = sum(!is.na(c_across(contains("abundance"))))) |>
  select(sequence, amount_quant, contains('abundance'))

mixed_database_identified_per_peptide <- mixed_database_peptides |>
  filter(!if_all(contains('Abundance'), ~ is.na(.))) |>
  mutate(Sequence = str_split_i(`Annotated Sequence`,'\\.',2)) |>
  select(Sequence, contains('Abundance:')) |>
  rowwise() |>
  mutate(amount_quant = sum(!is.na(c_across(contains("Abundance"))))) |>
  select(Sequence, amount_quant, contains('abundance'))

```

```{r}

cat(str_c("Mean amount of quantification values per peptide\nINQuant: ",
      mean(mixed_identified_per_peptide$amount_quant),
      "\nDatabase: ",
      mean(mixed_database_identified_per_peptide$amount_quant)))

```

#### INQuant and database search protein identification Venn diagram

```{r, output = FALSE}

mixed_inquant_proteins_identified <- mixed_MBR_proteins |>
  pull(principal_protein)

mixed_database_proteins_identified <- mixed_database_proteins |>
  pull(Accession)

mixed_venn_proteins <- plot_venn_diagram('INQuant',
                  'Database',
                  mixed_inquant_proteins_identified,
                  mixed_database_proteins_identified) +
  plot_annotation(title = 'Indentified proteins in the mixed proteome experiment',
                  theme = theme(plot.title = element_text(hjust = 0.5),
                                plot.subtitle = element_text(hjust = 0.5)))

```

```{r}

save_fig(mixed_venn_proteins, 
         'mixed_venn_identified_proteins.png', 
         w = 1600,
         h = 1600)

```

### The GluC experiment

#### Density scatter plot for calibration

```{r, output = FALSE}

plot_meas_rt <- plot_density_point(gluc_calibrated_preds, "rt", "Measured", 7)
plot_cali_rt <- plot_density_point(gluc_calibrated_preds, "rt", "Calibrated", 7)
plot_meas_mz <- plot_density_point(gluc_calibrated_preds, "mz", "Measured", 7)
plot_cali_mz <- plot_density_point(gluc_calibrated_preds, "mz", "Calibrated", 7)

calibration_scatter_error_plot <- 
  (plot_meas_rt + 
     plot_cali_rt + 
     plot_meas_mz + 
     plot_cali_mz) + 
  plot_layout(ncol = 2, guides = "collect") + 
  plot_annotation(title = "Calibration density plots",
                  theme = theme(plot.title = element_text(hjust = 0.5), 
                                legend.position = "bottom"))

```

```{r}

save_fig(calibration_scatter_error_plot, 
         "calibration_scatter_error_plot.png")

```

#### Density graph

```{r, warning = FALSE}

density_cali <- plot_density_calibration(gluc_calibrated_preds, calibrated = TRUE)
density_not_cali <- plot_density_calibration(gluc_calibrated_preds, calibrated = FALSE)

save_fig(
  (density_not_cali + theme(legend.position = "none")) / 
    (density_cali + theme(legend.position = "bottom")), 
  "calibration_density_error_plot_cali.png", 
  w = 1920/2
)
```

#### Calibration m/z error boxplot

```{r, warning = FALSE}

boxplot_cali_outliers <- plot_boxplot_calibration(gluc_calibrated_preds, 
                             outliers = TRUE)
save_fig(boxplot_cali_outliers,
         "calibration_boxplot.png", 
         w = 1920/2)

```

#### Distribution of C-terminal amino acids

```{r, output = FALSE}

gluc_cleavage_distribution <- 
  plot_c_term_AA_distribution(gluc_peptides, 
                              'INQuant', 
                              'abundance_(.*)')

gluc_database_cleavage_distribution <- 
  plot_c_term_AA_distribution(
    gluc_database_peptides |>
      mutate(sequence = Sequence), 
  'Database', 
  'Abundances \\(Normalized\\): (.*): Sample')

gluc_cleavage_combined <- 
  (gluc_cleavage_distribution + gluc_database_cleavage_distribution +
     plot_layout(2,1,axis_titles = 'collect', guides = 'collect') + 
     plot_annotation(
       title = 
         'GluC: Peptide abundance percentage distribution for each experiment',
       subtitle = 
         'Categorised by C-terminal amino acid')
  )

```

```{r}

save_fig(gluc_cleavage_combined, 
              'GluC_database_inquant_C_term_cleavage_distribution.png',
              h = 1600)

```

#### Comparing MBR and non-MBR abundances

```{r}

save_fig(plot_bar_MBR_count(gluc_MBR_psms, 'GluC'), 
         'gluc_MBR_bar_plot.png', 
         h = 1340)

```

#### Identified quantification values distribution

```{r}

gluc_quant_amount <- full_join(gluc_peptides |>
            select(sequence, contains('abundance')) |>
            mutate(quantified = !if_all(contains('abundance'), is.na)) |>
            mutate(sequence = str_c(sequence, "_", row_number()), .by = sequence) |>
            select(sequence, quantified), 
          gluc_database_peptides |>
            select(`Sequence`, 
                   contains('ormalized')) |>
            mutate(sequence = str_replace_all(Sequence, 'I','L'),
                   quantified = !if_all(contains('Abundance'), is.na)) |>
            mutate(sequence = str_c(sequence, "_", row_number()), .by = sequence) |>
            select(sequence, quantified), 
          suffix = c('_INQuant', '_Database'),
          by = join_by(sequence)) |>
  pivot_longer(contains('quant'),
               names_to = 'type',
               values_to = 'quantified',
               names_prefix = 'quantified_') |>
  mutate(quantified = ifelse(quantified == 'TRUE', 'Quantified', 'Not quantified')) |>
  mutate(quantified = replace_na(quantified, 'Not identified')) |>
  count(type, quantified) |>
  group_by(type) |>
  mutate(percentage = n / sum(n),
         label = str_c(round(percentage*100, digits = 1),'% \n(',n,')')) |>
  ungroup() |>
  ggplot(aes(x = type, y = percentage, fill = quantified)) +
  geom_col(width = 0.8, alpha = 0.8) +
  geom_label(aes(label = label),
            position = position_fill(vjust = 0.5),
            alpha = 0.6,
            label.size = 0,
            size = 3, 
            color = "white",
            show.legend = FALSE) +
  scale_fill_manual(values = c('Quantified' = inq_blue,
                               'Not quantified' = inq_pink,
                               'Not identified' = 'grey')) +
  coord_flip() +
  theme_minimal() +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(title = 'GluC: All identified peptides',
       subtitle = 'Categorised by whether or not they have quantification values',
       fill = '',
       x = '',
       y = '')

save_fig(gluc_quant_amount, 
         'GluC_amount_quantified_database_vs_INQuant.png',
              h = 1200,
              w = 2300)

```

#### Identified peptides Venn diagram

```{r, output = FALSE}

gluc_identified <- gluc_MBR_peptides |>
  filter(!if_all(contains('abundance'), ~ is.na(.))) |>
  mutate(sequence = str_c(sequence, "_", row_number()), .by = sequence) |>
  pull(sequence)

gluc_database_identified <- gluc_database_peptides |>
  filter(!if_all(contains('Abundance'), ~ is.na(.))) |>
  mutate(Sequence = str_replace_all(Sequence, 'I','L')) |>
  mutate(sequence = str_c(Sequence, "_", row_number()), .by = Sequence) |>
  pull(sequence)

gluc_venn <- plot_venn_diagram('INQuant', 'Database',
                  gluc_identified,
                  gluc_database_identified) +
  plot_annotation(title = 'Quantified peptides in the GluC experiment',
                  subtitle = 'With Database Is replaced by Ls',
                  theme = theme(plot.title = element_text(hjust = 0.5),
                                plot.subtitle = element_text(hjust = 0.5)))

```

```{r}

save_fig(gluc_venn, 'GluC_venn_identified_peptides.png', 
              w = 1600,
              h = 1600)

```

#### Fold change boxplots

```{r}

gluc_fold_change_inquant_data_with_MBR <- gluc_MBR_proteins |>
  select(!matches('abundance_(Ctrl\\d|GluC\\d)$')) |>
  pivot_longer(cols = contains('abundance'),
               names_to = c("experiment"),
               names_pattern = "abundance_(.*)\\d_normalized",
               values_to = "abundance_normalized") |>
  drop_na(abundance_normalized) |>
  group_by(principal_protein, experiment) |>
  summarise(mean_abundance_normalized = mean(abundance_normalized), .groups = 'drop') |>
  pivot_wider(names_from = experiment,
              values_from = mean_abundance_normalized) |>
  drop_na() |>
  mutate(fold_change = Ctrl / GluC,
         origin = "INQuant (with MBR)") |>
  select(fold_change, origin)

gluc_fold_change_inquant_data_without_MBR <- gluc_proteins |>
  select(!matches('abundance_(Ctrl\\d|GluC\\d)$')) |>
  pivot_longer(cols = contains('abundance'),
               names_to = c("experiment"),
               names_pattern = "abundance_(.*)\\d_normalized",
               values_to = "abundance_normalized") |>
  drop_na(abundance_normalized) |>
  group_by(principal_protein, experiment) |>
  summarise(mean_abundance_normalized = mean(abundance_normalized), .groups = 'drop') |>
  pivot_wider(names_from = experiment,
              values_from = mean_abundance_normalized) |>
  drop_na() |>
  mutate(fold_change = Ctrl / GluC,
         origin = "INQuant (without MBR)") |>
  select(fold_change, origin)

gluc_fold_change_database_data <- gluc_database_proteins |>
  select(!contains('Abundance:') & !contains('Scaled')) |>
  pivot_longer(cols = contains('Abundances'),
               names_to = c("experiment"),
               names_pattern = "Abundances \\(Normalized\\): (.*)\\d: Sample",
               values_to = "abundance_normalized") |>
  drop_na(abundance_normalized) |>
  group_by(Accession, experiment) |>
  summarise(mean_abundance_normalized = mean(abundance_normalized), .groups = 'drop') |>
  pivot_wider(names_from = experiment,
              values_from = mean_abundance_normalized) |>
  drop_na() |>
  mutate(fold_change = Ctrl / GluC,
         origin = "Database") |>
  select(fold_change, origin)

gluc_fold_change_boxplot <- 
  bind_rows(gluc_fold_change_inquant_data_with_MBR, 
            gluc_fold_change_database_data) |>
  mutate(
    log2_fc = log2(fold_change),
    origin = factor(origin, levels = c("INQuant (with MBR)", 
                                       "Database"))  
  ) |>
  ggplot(aes(x = origin, y = log2_fc, fill = origin)) +  
  geom_boxplot(alpha = 0.8, outlier.alpha = 0.6) +
  scale_fill_manual(
    values = c("INQuant (with MBR)" = inq_pink, "Database" = inq_blue)
  ) + 
  geom_hline(aes(yintercept = 0), 
             linetype = 'dashed', 
             color = inq_purple,
             show.legend = FALSE) +
  labs(
    y = "log2(fold change)",
    x = NULL,
    fill = "Data origin",
    title = "GluC: Log2 fold change of normalized protein abundance"
  ) +
  theme_minimal() + 
  theme(legend.position = "none", 
        legend.direction = "horizontal",
        title = element_text(size = 10))

save_fig(gluc_fold_change_boxplot, 'gluc_fold_change_boxplot.png', w = 1920*4/5)

gluc_fold_change_boxplot_MBR_diff <- 
  bind_rows(gluc_fold_change_inquant_data_with_MBR, 
            gluc_fold_change_inquant_data_without_MBR,
            gluc_fold_change_database_data) |>
  mutate(
    log2_fc = log2(fold_change),
    origin = factor(origin, levels = c("INQuant (without MBR)", 
                                       "INQuant (with MBR)", 
                                       "Database"))  
  ) |>
  ggplot(aes(x = origin, y = log2_fc, fill = origin)) +  # Explicit x = origin for clarity
  geom_boxplot(alpha = 0.8, outlier.alpha = 0.6) +
  scale_fill_manual(
    values = c("INQuant (with MBR)" = inq_pink,
               "INQuant (without MBR)" = inq_orange, 
               "Database" = inq_blue)
  ) + 
  geom_hline(aes(yintercept = 0), 
             linetype = 'dashed', 
             color = inq_purple,
             show.legend = FALSE) +
  labs(
    y = "log2(fold_change)",
    x = NULL,
    fill = "Data origin",
    title = "GluC: Log2 fold change of normalized protein abundance",
    subtitle = "For both INQuant and Database search data"
  ) +
  theme_minimal() 
  
save_fig(gluc_fold_change_boxplot_MBR_diff, 'gluc_fold_change_boxplot_MBR_diff.png')

save_fig(gluc_fold_change_boxplot_MBR_diff + 
  coord_cartesian(ylim = c(-1,1)), 'gluc_fold_change_boxplot_MBR_diff_zoomed.png')

```

#### High confidence peptides without alignment to proteins

```{r}

cat(str_c('Number of peptides with a confidence 95%: ',
gluc_MBR_peptides |> 
  filter(protein_matches == '') |> 
  mutate(conf_perc = 10**conf*100) |> 
  filter(conf_perc >= 0.95) |>
  count() |>
  pull()
))

```

#### Identified proteins venn diagram

```{r, output = FALSE}

gluc_inquant_proteins_identified <- gluc_MBR_proteins |>
  pull(principal_protein)

gluc_database_proteins_identified <- gluc_database_proteins |>
  pull(Accession)

gluc_venn_proteins <- 
  plot_venn_diagram('INQuant',
                  'Database',
                  gluc_inquant_proteins_identified,
                  gluc_database_proteins_identified) +
  plot_annotation(title = 'Indentified proteins in the GluC experiment',
                  theme = theme(plot.title = element_text(hjust = 0.5),
                                plot.subtitle = element_text(hjust = 0.5)))

```

```{r}

save_fig(gluc_venn_proteins, 'gluc_venn_identified_proteins.png', 
              w = 1600,
              h = 1600)

```

#### Amount of correct and wrong predictions before and after recalibration

```{r}

combined_calibrated_preds <- gluc_calibrated_preds |> 
  full_join(gluc_recalibrated_preds, by = c("file", "scan_number")) |> 
  filter(!is.na(target)) |>
  mutate(best = ifelse(target == preds, 
                       ifelse(target == preds_cali, "both", "preds"),
                       ifelse(target == preds_cali, "cali", "none")), 
         best = ifelse(is.na(best), 
                       ifelse(is.na(preds), 
                              ifelse(is.na(preds_cali), "none", 
                                     ifelse(target == preds_cali, "cali", "none")),
                              ifelse(is.na(preds_cali), 
                                     ifelse(target == preds, "preds", "none"), 
                                     best)),
                       best)) |> 
  select(!global_index, !correct)

```

##### Before recalibration

```{r}

combined_calibrated_preds |> 
  filter(!is.na(target)) |>
  mutate('Prediction type' = ifelse(best == "both" | best == "preds", 'Correct', 'Wrong')) |>
  group_by(`Prediction type`) |>
  summarise(Peptides = n(), 
            'Mean confidence' = mean(probs_cali)) 


```

##### After recalibration

```{r}

combined_calibrated_preds |> 
  filter(!is.na(target)) |>
  mutate('Prediction type' = ifelse(best == "both" | best == "cali", 'Correct', 'Wrong')) |>
  group_by(`Prediction type`) |>
  summarise(Peptides = n(), 
            'Mean confidence' = mean(probs_cali)) 

```

#### Venn diagram before and after calibration

```{r}

gluc_identified_before <- combined_calibrated_preds |>
  filter(target == preds) |>
  mutate(target_scan = str_c(target, scan_number)) |>
  pull(target_scan)

gluc_identified_after <- combined_calibrated_preds |>
  filter(target == preds_cali) |>
  mutate(target_scan = str_c(target, scan_number)) |>
  pull(target_scan)
  
gluc_venn <- plot_venn_diagram('Before calibration', 'After calibration',
                  gluc_identified_before,
                  gluc_identified_after) +
  plot_annotation(title = 'Correctly identified peptides in the GluC experiment',
                  subtitle = 'Before and after calibration',
                  theme = theme(plot.title = element_text(hjust = 0.5),
                                plot.subtitle = element_text(hjust = 0.5)))

save_fig(gluc_venn, 'GluC_venn_calibration_identified_sequences.png', 
              w = 1600,
              h = 1600)
```

### NPv3

#### Boxplot for protein abundances

```{r}

npv3_boxplot <- NPv3_MBR_proteins |>
  mutate(protein_type = ifelse(str_detect(principal_protein,'^B_fig'),
                               'Native', 
                               ifelse(str_detect(principal_protein, '^P'), 
                                      'Phage', 'Anti-viral'))) |>
  select(!matches('abundance_.*\\d$')) |>
  pivot_longer(contains('abundance'),
               names_to = 'experiment',
               names_pattern = 'abundance\\_(.*)_normalized',
               values_to = 'abundance') |>
  mutate(experiment = ifelse(experiment == '40', '40_rep1', 
                             ifelse(experiment == 5, '05', experiment))) |>
  drop_na(abundance) |>
  ggplot() +
  geom_boxplot(aes(y = abundance, x = experiment, fill = protein_type), 
               outliers=FALSE,
               alpha = 0.7) + 
  scale_fill_manual(values = c('Anti-viral' = inq_purple,
                               'Native' = inq_dark_blue,
                               'Phage' = inq_orange
                               )) + 
  theme_minimal() + 
  labs(title = 'NPv3: Protein quantification for types of proteins with MBR',
       x = 'Experiment timestamp',
       y = 'Abundance',
       fill = 'Protein type')

save_fig(npv3_boxplot, 'NPv3_boxplot.png', h = 1440*0.6)

```

```{r}
NPv3_MBR_peptide |>  
  filter(str_detect(protein_matches, 'P_')) |> 
  summarise(Count = n(), 
            PSMs = sum(no_psms))
```

#### Protein inference, plot group for specific protein

```{r}

protein_group <- plot_groups(ungrouped_protein_table = NPv3_MBR_proteins_ungrouped,
            protein_group_table = NPv3_MBR_proteins,
            selected_protein = "P_ID000365",
            show_protein_names = TRUE)

save_fig(protein_group, 'NPv3_protein_group.png', 
         w = 1920*4/5)

```

```{r}

B_fig_peptide <- NPv3_MBR_proteins |> 
  filter(principal_protein == 'B_fig|287.41904.peg.5652') |>
  select(peptide_seqs) |> 
  str_split(';') |> 
  as_vector()

npv3_peptide_abund <- NPv3_MBR_peptide |>
  filter(sequence %in% B_fig_peptide) |>
  mutate(protein_type = case_when(
    str_detect(principal_protein, '^B_fig') ~ 'Native',
    str_detect(principal_protein, '^P') ~ 'Phage',
    TRUE ~ 'Anti-viral'
  )) |>
  select(!matches('abundance_.*\\d$')) |> 
  mutate(phage = str_detect(protein_matches, 'P_')) |> 
  pivot_longer(contains('abundance'),
               names_to = 'experiment',
               names_pattern = 'abundance\\_(.*)_normalized',
               values_to = 'abundance') |> 
  drop_na(abundance) |>
  mutate(experiment = case_when(
    experiment == '40' ~ '40_rep1',
    experiment == '5' ~ '05',
    TRUE ~ experiment
  )) |>
  drop_na(abundance) |>
  arrange(phage) |> 
  ggplot(aes(y = abundance, x = experiment, color = phage, shape = phage)) +
  geom_line(aes(group = seq_modifications)) +
  geom_point(aes(size = phage)) + 
  scale_color_manual(values = c("TRUE" = inq_orange, "FALSE" = inq_dark_blue),
                     labels = c("TRUE" = "Mapped to phage", "FALSE" = "Only mapped to native"),
                     guide = "legend") +
  scale_shape_manual(values = c(16, 17), 
                     labels = c("TRUE" = "Mapped to phage", "FALSE" = "Only mapped to native"),
                     guide = "legend") +
  scale_size_manual(values = c("TRUE" = 3, "FALSE" = 2), 
                     labels = c("TRUE" = "Mapped to phage", "FALSE" = "Only mapped to native"),
                     guide = "legend") +
  labs(
    title = "Abundance distribution across experiments",
    subtitle = "for peptides mapped to B_fig|287.41904.peg.5652",
    x = "Experiment",
    y = "Abundance",
    color = "Protein Mapping",
    shape = "Protein Mapping",
    size = "Protein Mapping"
  ) +
  theme_minimal() +
  theme(
    legend.position = 'right',
    legend.text = element_text(size = 8), 
    legend.title = element_text(size = 9),
    plot.title = element_text(size = 12),
    plot.subtitle = element_text(size = 11), 
    axis.title = element_text(size = 9),
  )


save_fig(npv3_peptide_abund, 'NPv3_peptide_abundance.png', 
         h = 1440*0.55)

```

```{r}
NPv3_MBR_peptide |> 
  filter(str_detect(principal_protein, '^P_')) |> 
  select(sequence, protein_matches, conf) |> 
  mutate(conf_perc = 10**conf)
```

## T-tests for MBR

### GluC

```{r}

gluc_paired_mean_abundances <- gluc_MBR_psms |>
  mutate(meas_mass = round(meas_mass, digits = 2),
         MBR = is.na(conf),
         file_id = str_extract(file_id, '\\D*')) |>
  group_by(sequence, meas_mass, MBR, file_id) |>
  summarise(mean_abundance = mean(abundance, na.rm = TRUE), .groups = 'drop') |>
  pivot_wider(names_from = MBR,
              names_prefix = 'mean_abundance_MBR_',
              values_from = mean_abundance) |>
  drop_na()

t.test(x = gluc_paired_mean_abundances$mean_abundance_MBR_FALSE,
       y = gluc_paired_mean_abundances$mean_abundance_MBR_TRUE,
       paired = TRUE)

```

### Mixed proteome

```{r}

mixed_paired_mean_abundances <- mixed_MBR_psms |>
  filter(file_id != 'A1') |>
  mutate(meas_mass = round(meas_mass, digits = 2),
         MBR = is.na(conf),
         file_id = str_extract(file_id, '\\D*')) |>
  group_by(sequence, meas_mass, MBR, file_id) |>
  summarise(mean_abundance = mean(abundance, na.rm = TRUE), .groups = 'keep') |>
  pivot_wider(names_from = MBR,
              names_prefix = 'mean_abundance_MBR_',
              values_from = mean_abundance) |>
  drop_na()

t.test(x = mixed_paired_mean_abundances$mean_abundance_MBR_FALSE,
       y = mixed_paired_mean_abundances$mean_abundance_MBR_TRUE,
       paired = TRUE)
```

### NPv3

```{r}

NPv3_paired_mean_abundances <- NPv3_MBR_psms |>
  filter(str_detect(file_id, '40')) |>
  mutate(meas_mass = round(meas_mass, digits = 2),
         MBR = is.na(conf)) |>
  group_by(sequence, meas_mass, MBR) |>
  summarise(mean_abundance = mean(abundance, na.rm = TRUE), .groups = 'drop') |>
  pivot_wider(names_from = MBR,
              names_prefix = 'mean_abundance_MBR_',
              values_from = mean_abundance) |>
  drop_na()

t.test(x = NPv3_paired_mean_abundances$mean_abundance_MBR_FALSE,
       y = NPv3_paired_mean_abundances$mean_abundance_MBR_TRUE,
       paired = TRUE)
```

### Average confidences

```{r}

combined_conf <- bind_rows(
  tibble(exp = 'GluC', conf = gluc_psms$conf),
  tibble(exp = 'Mixed proteome', conf = mixed_psms$conf),
  tibble(exp = 'NPv3', conf = NPv3_psms$conf)) |>
  drop_na() |>
  mutate(conf = 10**conf)

conf_medians <- combined_conf |>
  group_by(exp) |>
  summarise(median = median(conf))

combined_conf_plot <- 
  ggplot(combined_conf, aes(x = exp, y = conf, fill = exp, color = exp)) +
  geom_half_violin(
    aes(group = exp),
    side = "r", 
    alpha = 0.6,
    scale = 'width',
  ) +
  geom_boxplot(
    outlier.shape = 21,
    outlier.alpha = 0.5,
    outlier.size = 2,
    staplewidth = 0.3,
    size = .9,
    alpha = 0.6,
    width = 0.3,
    position = position_nudge(x = -0.2)
  ) +
  geom_label(data = conf_medians,
             aes(x = exp, 
                 y = median, 
                 label = format(median, digits = 1, scientific = FALSE), fill = exp),
             size = 3,
             nudge_x = -0.3,
             nudge_y = 0.05,
             color = 'white',
             label.size = 0,
             inherit.aes = FALSE) +
  scale_fill_manual(values = c(GluC = inq_dark_blue,
                               'Mixed proteome' = inq_purple,
                               NPv3 = inq_blue)) +
  scale_color_manual(values = c(GluC = inq_dark_blue,
                                'Mixed proteome' = inq_purple,
                                NPv3 = inq_blue)) +
  labs(title = 'Distribution of confidences for PSM identifications for each experiment',
       subtitle = 'For InstaNovo predictions',
       color = 'Experiment',
       x = '',
       y = 'Confidence') +
  theme_minimal() +
  theme(legend.position = 'none')

save_fig(combined_conf_plot, 'confidence_boxplots_all_exp.png')

```

### Confidence filter

```{r}

gluc_paired_mean_abundances_filtered <- gluc_MBR_psms |>
  mutate(meas_mass = round(meas_mass, digits = 2),
         MBR = is.na(conf),
         file_id = str_extract(file_id, '\\D*')) |>
  mutate(conf = ifelse(is.na(conf), 0, conf)) |>
  filter(10**conf >= 0.95) |>
  mutate(conf = ifelse(MBR, NA, conf)) |>
  group_by(sequence, meas_mass, MBR, file_id) |>
  summarise(mean_abundance = mean(abundance, na.rm = TRUE), .groups = 'drop') |>
  pivot_wider(names_from = MBR,
              names_prefix = 'mean_abundance_MBR_',
              values_from = mean_abundance) |>
  drop_na()

t.test(x = gluc_paired_mean_abundances_filtered$mean_abundance_MBR_FALSE,
       y = gluc_paired_mean_abundances_filtered$mean_abundance_MBR_TRUE,
       paired = TRUE)

```

```{r}

mixed_paired_mean_abundances_filtered <- mixed_MBR_psms |>
  filter(file_id != 'A1') |>
  mutate(meas_mass = round(meas_mass, digits = 2),
         MBR = is.na(conf),
         file_id = str_extract(file_id, '\\D*')) |>
  mutate(conf = ifelse(is.na(conf), 0, conf)) |>
  filter(10**conf >= 0.95) |>
  mutate(conf = ifelse(MBR, NA, conf)) |>
  group_by(sequence, meas_mass, MBR, file_id) |>
  summarise(mean_abundance = mean(abundance, na.rm = TRUE), .groups = 'keep') |>
  pivot_wider(names_from = MBR,
              names_prefix = 'mean_abundance_MBR_',
              values_from = mean_abundance) |>
  drop_na()

t.test(x = mixed_paired_mean_abundances_filtered$mean_abundance_MBR_FALSE,
       y = mixed_paired_mean_abundances_filtered$mean_abundance_MBR_TRUE,
       paired = TRUE)
```

```{r}

NPv3_paired_mean_abundances_filtered <- NPv3_MBR_psms |>
  filter(str_detect(file_id, '40')) |>
  mutate(meas_mass = round(meas_mass, digits = 2),
         MBR = is.na(conf)) |>
  mutate(conf = ifelse(is.na(conf), 0, conf)) |>
  filter(10**conf >= 0.95) |>
  mutate(conf = ifelse(MBR, NA, conf)) |>
  group_by(sequence, meas_mass, MBR) |>
  summarise(mean_abundance = mean(abundance, na.rm = TRUE), .groups = 'drop') |>
  pivot_wider(names_from = MBR,
              names_prefix = 'mean_abundance_MBR_',
              values_from = mean_abundance) |>
  drop_na()

t.test(x = NPv3_paired_mean_abundances_filtered$mean_abundance_MBR_FALSE,
       y = NPv3_paired_mean_abundances_filtered$mean_abundance_MBR_TRUE,
       paired = TRUE)
```
